Peer Assessment 1
========================================================

This are my answers to the questions posed in peer assessment 1.  
Loading of data

```{r loading of data}
temp = tempfile()
download.file("http://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip", temp)
data = read.csv(unz(temp, "activity.csv"), colClasses = c("numeric", "Date", "numeric"))
```

The mean number of steps made each day is:
```{r calculating mean NA ignored}
mean = tapply(data$steps,data$date, mean, na.rm = T)
print(mean)
```

The median number of steps made each day is:
```{r calculating median NA ignored}
median = tapply(data$steps,data$date, median, na.rm = T)
print(median)
```

This histogram shows the total amount of steps made each day:
```{r histogram step sum}
library(ggplot2)
x = tapply(data$steps, data$date, sum, na.rm = T)
x = as.data.frame(x)
g = ggplot(x, aes(rownames(x), x))
a = g + geom_bar(stat = "identity") + theme(axis.text.x = element_text(angle = 90, vjust = 0.5, size = 5)) +
        labs (x = "Day") + labs (y = "Number of Steps")
print (a)
```

The average daily pattern of steps is reflected in the next figure:
```{r average daily pattern}
x= tapply(data$steps, data$interval, mean, na.rm = T)
x1 =data.frame(as.numeric(rownames(x)), x)
g = ggplot(x1, aes(x1[,1], x1[,2]))
a = g + geom_line(stat = "identity") + theme(axis.text.x = element_text(angle = 90)) +
        labs (x = "Interval") + labs (y = "Number of Steps") + scale_x_discrete(breaks = seq(0, 2355, 50))
print (a)
````

 

```{r maximum interval}
maximum = x1[x1[, 2] == max(x1[,2]), 1]
```

With the maximum number of steps achieved at interval `r maximum`:

```{r amount missing values}
miss = nrow(data) - sum(complete.cases(data))
```

Data has `r miss` missing values, and missing values have been replaced by the average step number for each month in a new dataset. The following histogram shows the total number of steps taken each day after replacing the missing values.

```{r histogram after replace}
newdata = data.frame(data)
newdata$fecha = as.POSIXlt(data$date)
newdata[(!complete.cases(newdata[, 1]) & newdata$fecha$mon == 9), 1] = mean(newdata[newdata$fecha$mon== 9, 1], na.rm = T)
newdata[(!complete.cases(newdata[, 1]) & newdata$fecha$mon == 10), 1] = mean(newdata[newdata$fecha$mon== 10, 1], na.rm = T)
y = tapply(newdata$steps, newdata$date, sum, na.rm = T)
y = as.data.frame(y)
g = ggplot(y, aes(rownames(y), y))
a = g + geom_bar(stat = "identity") + theme(axis.text.x = element_text(angle = 90, vjust = 0.5, size = 5)) +
        labs (x = "Day") + labs (y = "Number of Steps")
print (a)
mean = tapply(newdata$steps,newdata$date, mean, na.rm = T)
median = tapply(newdata$steps,newdata$date, median, na.rm = T)
```

The mean of data has changed after the replacement.

```{r modified mean}
print(mean)
```
The median does so in days where all values were NA's
```{r modified median}
print (median)
```

The impact of this replacement on daily number of steps is that of rising the value for those days with NA's, particularly those days with no data at all.

Exploring the differences between weekdays and weekends, the following plot:

```{r weekdays and weekends}
a = c("sábado", "domingo")
for (i in 1:length(newdata$fecha)){
        if (weekdays(newdata$fecha[i]) %in% a){
                newdata$week[i] = "weekend"
        } else {
                newdata$week[i] = "weekday"
        }
}
m = aggregate (steps ~ interval + week, FUN = sum, data = newdata)
g = ggplot(m, aes(interval, steps))
a = g + geom_line(stat = "identity") + theme(axis.text.x = element_text(angle = 90)) +
        labs (x = "Interval") + labs (y = "Number of Steps") + scale_x_discrete(breaks = seq(0, 2355, 50)) + facet_grid(week~.)
print (a)
```

clearly shows that weekends are "lazier" than weekdays.
